\ \\ \LARGE\bf Speculative Evaluation in Particle Swarm Optimization \thanks{Matthew Gardner, Andrew McNabb, and Kevin Seppi are with the Department of Computer Science, Brigham Young University, 3361 TMCB, Provo, UT 84602 (phone: 801-422-8717; email: \{mjg82,a,k\}@cs.byu.edu).}

### 2010-03-22 honors_thesis/journal_paper/paper.tex
started notation and alg updates

%% PSO Stuff
\DeclareMathOperator*{\argmin}{arg\;min}
\DeclareMathOperator*{\argmax}{arg\;max}
\DeclareMathOperator*{\arginf}{arg\;inf}
\DeclareMathOperator*{\argsup}{arg\;sup}
\providecommand{\pers}{\ensuremath{P}}
\providecommand{\neigh}{\ensuremath{N}}
\providecommand{\nURand}{\ensuremath{U^\neigh}}
\providecommand{\pURand}{\ensuremath{U^\pers}}
\providecommand{\ppos}{\ensuremath{\Vec{x}}}
\providecommand{\fppos}{\ensuremath{\Vec{fx}}}
\providecommand{\pvel}{\ensuremath{\Vec{v}}}
\providecommand{\nbest}{\ensuremath{\Vec{b}^\neigh}}
\providecommand{\fnbest}{\ensuremath{\Vec{fb}^\neigh}}
\providecommand{\pbest}{\ensuremath{\Vec{b}^\pers}}
\providecommand{\fpbest}{\ensuremath{\Vec{fb}^\pers}}
\providecommand{\constriction}{\ensuremath{\chi}}
\providecommand{\ncoeff}{\ensuremath{\phi^\neigh}}
\providecommand{\pcoeff}{\ensuremath{\phi^\pers}}
\providecommand{\obs}{\ensuremath{\Vec{\xi}}}
\providecommand{\ofunc}{\ensuremath{f}}
\providecommand{\swarm}{\ensuremath{swarm}}

%SpecExPSO Stuff
\providecommand{\indic}{\ensuremath{I}}
\providecommand{\specvel}{\ensuremath{V}}
\providecommand{\specpos}{\ensuremath{X}}
\providecommand{\specval}{\ensuremath{FX}}
\providecommand{\ln}{\ensuremath{ln}}
\providecommand{\rn}{\ensuremath{rn}}
\providecommand{\caseset}{\ensuremath{\mathcal{C}}}
\providecommand{\casedef}{\ensuremath{(\pbest:\nbest)}}
\providecommand{\casexn}{\ensuremath{(x:\neigh*)}}
\providecommand{\casexx}{\ensuremath{(x:x)}}
\providecommand{\casexl}{\ensuremath{(x:\ln)}}
\providecommand{\casexr}{\ensuremath{(x:\rn)}}
\providecommand{\casepn}{\ensuremath{(\pers*:\neigh*)}}
\providecommand{\casepl}{\ensuremath{(\pers*:\ln)}}
\providecommand{\casepr}{\ensuremath{(\pers*:\rn)}}
\providecommand{\noeval}[1]{\ensuremath{#1^{-e}}}
\providecommand{\nonbest}[1]{\ensuremath{#1^{-n}}}
\providecommand{\p}{\ensuremath{p}}
\providecommand{\pset}{\ensuremath{\mathbf{p}}}
\providecommand{\s}{\ensuremath{s}}
\providecommand{\sset}{\ensuremath{\mathbf{s}}}
\providecommand{\nsset}{\ensuremath{\mathbf{ns}}}
\providecommand{\n}{\ensuremath{n}}
\providecommand{\nset}{\ensuremath{\mathbf{n}}}
\providecommand{\nnset}{\ensuremath{\mathbf{nn}}}

@@

At each iteration of the algorithm, the position $\ppos_t$ and velocity
$\pvel_t$ of each particle $i$ are updated as follows:
\begin{align}
\label{eq:velupdate}
	\pvel_{i,t+1} &=
		\constriction \left[ \pvel_t +
			\pcoeff\pURand_{i,t}\otimes(\pbest_{i,t} - \ppos_{i,t}) +
			\ncoeff\nURand_{i,t}\otimes(\nbest_{i,t} - \ppos_{i,t})
		\right] \\
\label{eq:posupdate}
	\ppos_{i,t+1} &= \ppos_{i,t} + \pvel_{i,t+1}
\end{align}
where \( \nURand_{i,t} \) and \( \nURand_{i,t} \) are vectors of independent
and identically distributed random numbers drawn from a uniform
distribution, the \( \otimes \) operator is an element-wise vector
multiplication, $\pbest$ (called personal best) is the best position the
current particle has seen, and $\nbest$ (called neighborhood best) is the best
position the neighbors of the current particle have seen~\cite{bratton-sis07}.  \( \ncoeff \), \(
\pcoeff \), and \( \constriction \) are parameters with prescribed values
required to ensure convergence (2.05, 2.05, and .73,
respectively)~\cite{clerc-tec02}. Note also that we have made the particle identifier $i$
and the time step $t$ or $t+1$ explicit in the subscript and moved other qualifiers,
such as the personal $\pers$ and personal best $\pers*$ qualifiers to the superscript position.
Superscripts can be miss read as exponents, but they are qualifiers throughout this paper.

@@

%First consider formula for the \emph{next} iteration of PSO, which is identical
%to Equations~\eqref{eq:velupdate} and \eqref{eq:posupdate} except the time
%subscripts are incremented.  As noted above and
For simplicity in this
explanation, we restrict our discussion here to the case of a Ring topology
with two neighbors which we will call the ``right neighbor'' and ``left
neighbor'' neighbors identified for each particle $i$ as $\rn(i)$ and $\ln(i)$
respectively.

@@

For minimization, we can define this example indicator variable representing the case
where for particle $i$ will need to update $\pbest$ with the better
position $x_{i,t}$ for the next iteration ($t+1$), but $\nbest$ will not updated by:
\begin{align}
\nonumber
	\indic_{i,t+1}^{\casexn} & (\ofunc ( \ppos_{i,t+1} ) ,\ofunc(\ppos_{\ln(i),t+1}),\ofunc(\ppos_{\rn(i),t+1}) ,\ofunc(\pbest_{i,t}) ,\ofunc(\nbest_{i,t}))= \\
\nonumber
	& 1 \ if \  \ofunc(\ppos_{i,t+1}) > \ofunc(\pbest_{i,t}) \\
\nonumber
	& \quad \quad \ \wedge \ \ofunc(\nbest_{i,t}) > \ofunc(\ppos_{i,t+1}) \\
\nonumber
	& \quad \quad \ \wedge \ \ofunc(\nbest_{i,t}) > \ofunc(\ppos_{\ln(i),t+1}) \\
\nonumber
	& \quad \quad \ \wedge \ \ofunc(\nbest_{i,t}) > \ofunc(\ppos_{\rn(i),t+1}) \\
\label{eq:deficasexn}
	& 0 \ otherwise
\end{align}

@@

For each cases identified in Table~\ref{tab:evals} and identified by a
corresponding indicator variable defined above, we also provide notation for the \emph{next} velocity
update equation; once again using $\casexn$ as an example:
\begin{align}
\nonumber
	\specvel_{i,t+1}^{\casexn} & (\ppos_{i,t+1}, \ppos_{\ln(i),t+1}, \ppos_{\rn(i),t+1},
        \pbest_{i,t} , \nbest_{i,t}, \pURand_{i,t}, \nURand_{i,t})= \\
\label{eq:defvocasexn}
		& \constriction \left[ \pvel_t +
			\pcoeff\pURand_{i,t+1}\otimes(\pbest_{i,t+1} - \ppos_{i,t}+1) +
			\ncoeff\nURand_{i,t+1}\otimes(\nbest_{i,t+1} - \ppos_{i,t+1})
		\right]
\end{align}

@@

For each cases identified in
Table~\ref{tab:evals}
and identified by a corresponding indicator variable defined above,
we define also a position update equation with its compact form; once again using $\casexn$ as an example:
\begin{align}
\label{eq:defpcasexn}
	\specpos_{i,t+1}^{\casexn} & (\ppos_{i,t+1} ,\ppos_{\ln(i),t+1},\ppos_{\rn(i),t+1} ,\pbest_{i,t} ,\nbest_{i,t})= \\
\nonumber
	\specpos_{i,t+1}^{\casexn} & =  \ppos_{i,t} + \specvel_{i,t+1}^{\casexn}
\end{align}

Lastly, we also define notation for the value of the objective function at these points,
once again using $\casexn$ as an example:
\begin{align}
\label{eq:deffcasexn}
	\specval_{i,t+1}^{\casexn} & (\ppos_{i,t+1} ,\ppos_{\ln(i),t+1},\ppos_{\rn(i),t+1} ,\pbest_{i,t} ,\nbest_{i,t})= \\
\nonumber
	\specval_{i,t+1}^{\casexn} & =  \ofunc(\specval_{i,t+1}^{\casexn}
\end{align}

@@

The value of the objective function at $\ppos_{i,t+1}$ is given by:
\begin{align}
\label{eq:val2update}
	\ofunc (\ppos_{i,t+1}) = \sum_{c \in \caseset} \ \specval_{i,t+1}^{c} \ \indic_{i,t+1}^{c}
\end{align}
or, with $\specpos$ and $\indic$ written with the parameters which were omitted above:
\begin{align}
\nonumber
\ofunc & (\ppos_{i,t+1}) = \sum_{c \in \caseset} \\
\nonumber
& \ofunc(\specpos_{i,t+1}^{c}(\ppos_{i,t+1},\ppos_{\ln(i),t+1},\ppos_{\rn(i),t+1},
\pbest_{i,t},\nbest_{i,t},\pURand_{i,t}, \nURand_{i,t})) \\
\label{eq:val2updatelong}
& \indic_{i,t}^{c}(\ofunc ( \ppos_{i,t+1} ) ,\ofunc(\ppos_{\ln(i),t+1}),\ofunc(\ppos_{\rn(i),t+1}) ,\ofunc(\pbest_{i,t}) ,\ofunc(\nbest_{i,t}))
\end{align}

@@

\begin{algorithm} [h!]
  \caption{Standard PSO (minimizing)}
  \label{alg:standard}
  \begin{algorithmic}[1]
    \STATE t=0
    \STATE Initialize the swarm including
    \FOR{$i \in \swarm$}
      \STATE {$\ppos_{i,t} \leftarrow Random position$}
      \STATE {$\pvel_{i,t} \leftarrow Random velocities$}
      \STATE {$\fpbest_{i,t} \leftarrow Max Float$}
      \STATE {$\fnbest_{i,t} \leftarrow Max Float$}
    \ENDFOR
    \REPEAT
        \STATE $t=t+1$
        \FOR{$i \in \swarm$}
          \STATE {$\fppos_{i.t} \leftarrow \ofunc(\ppos_{i,t})$}
        \ENDFOR
        \FOR{$i \in \swarm$}
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \label{alg1:startifs}
          \STATE $\pbest_{i,t}\leftarrow \pbest_{i,t-1}$
          \STATE $\fpbest_{i,t}\leftarrow \fpbest_{i,t-1}$
          \IF {$\fppos_{i,t}<\fpbest_{i,t-1}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{i,t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{i,t}$
          \ENDIF
          \STATE $\nbest_{i,t}\leftarrow \nbest_{i,t-1}$
          \STATE $\fnbest_{i,t}\leftarrow \fnbest_{i,t-1}$
          \IF {$\fppos_{i,t}<\fnbest_{i,t}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{i,t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{i,t}$
          \ENDIF
          \IF {$\fppos_{\ln(i),t}<\fnbest_{i,t}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{\ln(i),t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{\ln(i),t}$
          \ENDIF
          \IF {$\ofunc(\ppos_{i,t})<\ofunc(\nbest_{\rn(i),t})$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{\rn(i),t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{\rn(i),t}$
          \label{alg1:endifs}
          \ENDIF
          \STATE Update $\pvel_{i,t+1}$ as shown in equation~\eqref{eq:velupdate} 
          \STATE Update $\ppos_{i,t+1}$ as shown in equation~\eqref{eq:posupdate}
        \ENDFOR
    \UNTIL {$t >$ max iteration or convergence criteria met}
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm} [h!]
  \caption{Unrolled PSO}
  \label{alg:unrolled}
  \begin{algorithmic}[1]
    \STATE t=0
    \STATE Initialize the swarm including
    \FOR{$i \in \swarm$}
      \STATE {$\ppos_{i,t} \leftarrow Random position$}
      \STATE {$\pvel_{i,t} \leftarrow Random velocities$}
      \STATE {$\fpbest_{i,t} \leftarrow Max Float$}
      \STATE {$\fnbest_{i,t} \leftarrow Max Float$}
    \ENDFOR
    \REPEAT
        \STATE $t=t+1$
        \FOR{$i \in \swarm$}
          \label{alg2:fatitert}
          \STATE DISPATCH {$\fppos_{i.t} \leftarrow \ofunc(\ppos_{i,t})$} to a seperate processor
        \ENDFOR
        \FOR{$i \in \swarm$}
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \FOR{$c \in \caseset$}
            \STATE Compute $\specvel_{i,t+1}^c$ and $\specpos_{i,t+1}^c$.
	    Examples shown in equations~\eqref{eq:defvcasexn} and~\eqref{eq:defpcasexn}
            \label{alg2:fatitertp1}
            \STATE DISPATCH computation of $\specval_{i,t+1}^c$ to a seperate processor.
	    Example shown in equation~\eqref{eq:deffcasexn}.
	    \COMMENT Note that none of the computations, $\specvel_{i,t+1}^c$,
	    $\specpos_{i,t+1}^c$ and $\specval_{i,t+1}^c$ depend on $\fppos_{i.t}$.
          \ENDFOR
        \ENDFOR
        \STATE WAIT for all function evaluations dispatched above to complete
        \FOR{$i \in \swarm$}
          \FOR{$c \in \caseset$}
            \STATE Compute $\indic_{i,t+1}^c$
	           Examples shown in equation~\eqref{eq:deficasexn} and~\eqref{eq:defpcasexn}
	  \ENDFOR 
          \STATE Compute $\pvel_{i,t+1}$ as shown in equation~\eqref{eq:vel2update} 
          \STATE Compute $\ppos_{i,t+1}$ as shown in equation~\eqref{eq:pos2update}
          \STATE Compute $\fppos_{i,t+1}$ as shown in equation~\eqref{eq:val2update}
        \ENDFOR

        \STATE $t=t+1$
        \FOR{$i \in \swarm$}
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE $\pbest_{i,t}\leftarrow \pbest_{i,t-1}$
          \STATE $\fpbest_{i,t}\leftarrow \fpbest_{i,t-1}$
          \IF {$\fppos_{i,t}<\fpbest_{i,t-1}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{i,t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{i,t}$
          \ENDIF
          \STATE $\nbest_{i,t}\leftarrow \nbest_{i,t-1}$
          \STATE $\fnbest_{i,t}\leftarrow \fnbest_{i,t-1}$
          \IF {$\fppos_{i,t}<\fnbest_{i,t}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{i,t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{i,t}$
          \ENDIF
          \IF {$\fppos_{\ln(i),t}<\fnbest_{i,t}$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{\ln(i),t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{\ln(i),t}$
          \ENDIF
          \IF {$\ofunc(\ppos_{i,t})<\ofunc(\nbest_{\rn(i),t})$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{\rn(i),t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{\rn(i),t}$
          \ENDIF
          \STATE Update $\pvel_{i,t+1}$ as shown in equation~\eqref{eq:velupdate} 
          \STATE Update $\ppos_{i,t+1}$ as shown in equation~\eqref{eq:posupdate}
        \ENDFOR
    \UNTIL {$t >$ max iteration or convergence criteria met}
  \end{algorithmic}
\end{algorithm}

>>>>

%% PSO Stuff
\DeclareMathOperator*{\argmin}{arg\;min}
\DeclareMathOperator*{\argmax}{arg\;max}
\DeclareMathOperator*{\arginf}{arg\;inf}
\DeclareMathOperator*{\argsup}{arg\;sup}
\providecommand{\pers}{\ensuremath{P}}
\providecommand{\neigh}{\ensuremath{N}}
\providecommand{\leftind}{\ensuremath{L}}
\providecommand{\rightind}{\ensuremath{R}}
\providecommand{\nURand}{\ensuremath{U^\neigh}}
\providecommand{\pURand}{\ensuremath{U^\pers}}
\providecommand{\ppos}{\ensuremath{\Vec{x}}}
\providecommand{\fppos}{\ensuremath{\Vec{fx}}}
\providecommand{\pvel}{\ensuremath{\Vec{v}}}
\providecommand{\nbest}{\ensuremath{\Vec{b}^\neigh}}
\providecommand{\fnbest}{\ensuremath{\Vec{fb}^\neigh}}
\providecommand{\pbest}{\ensuremath{\Vec{b}^\pers}}
\providecommand{\fpbest}{\ensuremath{\Vec{fb}^\pers}}
\providecommand{\constriction}{\ensuremath{\chi}}
\providecommand{\ncoeff}{\ensuremath{\phi^\neigh}}
\providecommand{\pcoeff}{\ensuremath{\phi^\pers}}
\providecommand{\obs}{\ensuremath{\Vec{\xi}}}
\providecommand{\ofunc}{\ensuremath{f}}
\providecommand{\swarm}{\ensuremath{swarm}}

%SpecExPSO Stuff
\providecommand{\indic}{\ensuremath{I}}
\providecommand{\specvel}{\ensuremath{V}}
\providecommand{\specpos}{\ensuremath{X}}
\providecommand{\specval}{\ensuremath{FX}}
\providecommand{\leftn}{\ensuremath{\Vec{x}^\leftind}}
\providecommand{\rightn}{\ensuremath{\Vec{x}^\rightind}}
\providecommand{\caseset}{\ensuremath{\mathcal{C}}}
\providecommand{\casedef}{\ensuremath{(\pbest:\nbest)}}
\providecommand{\casexn}{\ensuremath{(x:\neigh)}}
\providecommand{\casexx}{\ensuremath{(x:x)}}
\providecommand{\casexl}{\ensuremath{(x:\leftind)}}
\providecommand{\casexr}{\ensuremath{(x:\rightind)}}
\providecommand{\casepn}{\ensuremath{(\pers:\neigh)}}
\providecommand{\casepl}{\ensuremath{(\pers:\leftind)}}
\providecommand{\casepr}{\ensuremath{(\pers:\rightind)}}
\providecommand{\noeval}[1]{\ensuremath{#1^{-e}}}
\providecommand{\nonbest}[1]{\ensuremath{#1^{-n}}}
\providecommand{\p}{\ensuremath{p}}
\providecommand{\pset}{\ensuremath{\mathbf{p}}}
\providecommand{\s}{\ensuremath{s}}
\providecommand{\sset}{\ensuremath{\mathbf{s}}}
\providecommand{\nsset}{\ensuremath{\mathbf{ns}}}
\providecommand{\n}{\ensuremath{n}}
\providecommand{\nset}{\ensuremath{\mathbf{n}}}
\providecommand{\nnset}{\ensuremath{\mathbf{nn}}}

@@

At each iteration of the algorithm, the position $\ppos_t$ and velocity
$\pvel_t$ of each particle $i$ are updated as follows:
\begin{align}
\nonumber
	\pvel_{i,t+1} &=
		\constriction \bigl[ \pvel_t
			+ \pcoeff\pURand_{i,t}\otimes(\pbest_{i,t} - \ppos_{i,t}) \\
\label{eq:velupdate}
			& \quad \quad + \ncoeff\nURand_{i,t}\otimes(\nbest_{i,t} - \ppos_{i,t})
		\bigr] \\
\label{eq:posupdate}
	\ppos_{i,t+1} &= \ppos_{i,t} + \pvel_{i,t+1}
\end{align}
where \( \nURand_{i,t} \) and \( \nURand_{i,t} \) are vectors of independent
and identically distributed random numbers drawn from a uniform
distribution, the \( \otimes \) operator is an element-wise vector
multiplication, $\pbest$ (called personal best) is the best position the
current particle has seen, and $\nbest$ (called neighborhood best) is the best
position the neighbors of the current particle have seen~\cite{bratton-sis07}.  \( \ncoeff \), \(
\pcoeff \), and \( \constriction \) are parameters with prescribed values
required to ensure convergence (2.05, 2.05, and .73,
respectively)~\cite{clerc-tec02}. Note also that we have made the particle identifier $i$
and the time step $t$ or $t+1$ explicit in the subscript and moved other qualifiers,
such as the personal $\pers$ and personal best $\pers*$ qualifiers to the superscript position.
Superscripts can be miss read as exponents, but they are qualifiers throughout this paper.

@@

%First consider formula for the \emph{next} iteration of PSO, which is identical
%to Equations~\eqref{eq:velupdate} and \eqref{eq:posupdate} except the time
%subscripts are incremented.  As noted above and
For simplicity in this
explanation, we restrict our discussion here to the case of a Ring topology
with two neighbors which we will call the ``right neighbor'' and ``left
neighbor'' neighbors identified for each particle $i$ as $\rightn$ and $\leftn$
respectively.

@@

For minimization, we can define this example indicator variable representing the case
where for particle $i$ will need to update $\pbest$ with the better
position $x_{i,t}$ for the next iteration ($t+1$), but $\nbest$ will not updated by:
\begin{align}
\nonumber
	\indic_{i,t+1}^{\casexn} & (\ofunc ( \ppos_{i,t+1} ) ,\ofunc(\leftn_{i,t+1}),\ofunc(\rightn_{i,t+1}) ,\ofunc(\pbest_{i,t}) ,\ofunc(\nbest_{i,t}))= \\
\nonumber
	& 1 \ if \  \ofunc(\ppos_{i,t+1}) > \ofunc(\pbest_{i,t}) \\
\nonumber
	& \quad \quad \ \wedge \ \ofunc(\nbest_{i,t}) > \ofunc(\ppos_{i,t+1}) \\
\nonumber
	& \quad \quad \ \wedge \ \ofunc(\nbest_{i,t}) > \ofunc(\leftn_{i,t+1}) \\
\nonumber
	& \quad \quad \ \wedge \ \ofunc(\nbest_{i,t}) > \ofunc(\rightn_{i,t+1}) \\
\label{eq:deficasexn}
	& 0 \ otherwise
\end{align}

@@

For each cases identified in Table~\ref{tab:evals} and identified by a
corresponding indicator variable defined above, we also provide notation for the \emph{next} velocity
update equation; once again using $\casexn$ as an example:
\begin{align}
\nonumber
	\specvel_{i,t+1}^{\casexn} & (\ppos_{i,t+1}, \leftn_{i,t+1}, \rightn_{i,t+1},
        \pbest_{i,t} , \nbest_{i,t}, \pURand_{i,t}, \nURand_{i,t})= \\
\label{eq:defvocasexn}
		& \constriction \left[ \pvel_t +
			\pcoeff\pURand_{i,t+1}\otimes(\pbest_{i,t+1} - \ppos_{i,t}+1) +
			\ncoeff\nURand_{i,t+1}\otimes(\nbest_{i,t+1} - \ppos_{i,t+1})
		\right]
\end{align}

@@

For each cases identified in
Table~\ref{tab:evals}
and identified by a corresponding indicator variable defined above,
we define also a position update equation with its compact form; once again using $\casexn$ as an example:
\begin{align}
\label{eq:defpcasexn}
	\specpos_{i,t+1}^{\casexn} & (\ppos_{i,t+1} ,\leftn_{i,t+1},\rightn_{i,t+1} ,\pbest_{i,t} ,\nbest_{i,t})= \\
\nonumber
	\specpos_{i,t+1}^{\casexn} & =  \ppos_{i,t} + \specvel_{i,t+1}^{\casexn}
\end{align}

Lastly, we also define notation for the value of the objective function at these points,
once again using $\casexn$ as an example:
\begin{align}
\label{eq:deffcasexn}
	\specval_{i,t+1}^{\casexn} & (\ppos_{i,t+1} ,\leftn_{i,t+1},\rightn_{i,t+1} ,\pbest_{i,t} ,\nbest_{i,t})= \\
\nonumber
	\specval_{i,t+1}^{\casexn} & =  \ofunc(\specval_{i,t+1}^{\casexn}
\end{align}

@@

The value of the objective function at $\ppos_{i,t+1}$ is given by:
\begin{align}
\label{eq:val2update}
	\ofunc (\ppos_{i,t+1}) = \sum_{c \in \caseset} \ \specval_{i,t+1}^{c} \ \indic_{i,t+1}^{c}
\end{align}
or, with $\specpos$ and $\indic$ written with the parameters which were omitted above:
\begin{align}
\nonumber
\ofunc & (\ppos_{i,t+1}) = \sum_{c \in \caseset} \\
\nonumber
& \ofunc(\specpos_{i,t+1}^{c}(\ppos_{i,t+1},\leftn_{i,t+1},\rightn_{i,t+1},
\pbest_{i,t},\nbest_{i,t},\pURand_{i,t}, \nURand_{i,t})) \\
\label{eq:val2updatelong}
& \indic_{i,t}^{c}(\ofunc ( \ppos_{i,t+1} ) ,\ofunc(\leftn_{i,t+1}),\ofunc(\rightn_{i,t+1}) ,\ofunc(\pbest_{i,t}) ,\ofunc(\nbest_{i,t}))
\end{align}

@@

First we define a helper function to update the personal and neighborhood bests, $\pbest$ and $\nbest$.
This algorithm is shown in \alg{alg:bests}.

\begin{algorithm}
  \caption{``UpdateBests'': Update personal and neighborhood best values (minimizing)}
  \label{alg:bests}
  \begin{algorithmic}[1]
          \STATE $\pbest_{i,t}\leftarrow \pbest_{i,t-1}$
          \STATE $\ofunc(\pbest_{i,t})\leftarrow \ofunc(\pbest_{i,t-1})$
          \IF {$\ofunc(\ppos_{i,t})<\ofunc(\pbest_{i,t-1})$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{i,t}$
            \STATE $\fpbest_{i,t}\leftarrow \fppos_{i,t}$
          \ENDIF
          \STATE $\nbest_{i,t}\leftarrow \nbest_{i,t-1}$
          \STATE $\ofunc(\nbest_{i,t})\leftarrow \ofunc(\nbest_{i,t-1})$
          \IF {$\ofunc(\fppos_{i,t})<\ofunc(\nbest_{i,t})$}
            \STATE $\pbest_{i,t}\leftarrow \ppos_{i,t}$
            \STATE $\ofunc(\pbest_{i,t})\leftarrow \ofunc(\ppos_{i,t})$
          \ENDIF
          \IF {$\ofunc(\leftn_{i,t})<\ofunc(nbest_{i,t})$}
            \STATE $\pbest_{i,t}\leftarrow \leftn_{i,t}$
            \STATE $\ofunc(pbest_{i,t})\leftarrow \ofunc(\leftn_{i,t})$
          \ENDIF
          \IF {$\ofunc(\rightn_{i,t})<\ofunc(\nbest_{i,t})$}
            \STATE $\pbest_{i,t}\leftarrow \rightn_{i,t}$
            \STATE $\ofunc(\pbest_{i,t})\leftarrow \ofunc(\rightn_{i,t}$
          \label{alg1:endifs}
          \ENDIF
          \STATE Update $\pvel_{i,t+1}$ as shown in equation~\eqref{eq:velupdate} 
          \STATE Update $\ppos_{i,t+1}$ as shown in equation~\eqref{eq:posupdate}
  \end{algorithmic}
\end{algorithm}

\begin{algorithm} [h!]
  \caption{Standard PSO (minimizing)}
  \label{alg:standard}
  \begin{algorithmic}[1]
    \STATE t=0
    \STATE Initialize the swarm
%    \FOR{$i \in \swarm$}
%      \STATE {$\ppos_{i,t} \leftarrow Random position$}
%      \STATE {$\pvel_{i,t} \leftarrow Random velocities$}
%      \STATE {$\fpbest_{i,t} \leftarrow Max Float$}
%      \STATE {$\fnbest_{i,t} \leftarrow Max Float$}
%    \ENDFOR
    \REPEAT
        \STATE $t=t+1$
        \FOR{$i \in \swarm$}
          \STATE {$\fppos_{i.t} \leftarrow \ofunc(\ppos_{i,t})$}
        \ENDFOR
        \FOR{$i \in \swarm$}
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE UpdateBests
          \STATE Update $\pvel_{i,t+1}$ as shown in equation~\eqref{eq:velupdate} 
          \STATE Update $\ppos_{i,t+1}$ as shown in equation~\eqref{eq:posupdate}
        \ENDFOR
    \UNTIL {$t > max \ iterations$ or convergence criteria met}
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm} [h!]
  \caption{Unrolled PSO}
  \label{alg:unrolled}
  \begin{algorithmic}[1]
    \STATE t=0
    \STATE Initialize the swarm including
%    \FOR{$i \in \swarm$}
%      \STATE {$\ppos_{i,t} \leftarrow Random position$}
%      \STATE {$\pvel_{i,t} \leftarrow Random velocities$}
%      \STATE {$\fpbest_{i,t} \leftarrow Max Float$}
%      \STATE {$\fnbest_{i,t} \leftarrow Max Float$}
%    \ENDFOR
    \REPEAT
        \STATE $t=t+1$
        \FOR{$i \in \swarm$}
          \label{alg2:fatitert}
          \STATE DISPATCH the computation of {$\ofunc(\ppos_{i.t})$} to a seperate processor
        \ENDFOR
        \FOR{$i \in \swarm$}
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \FOR{$c \in \caseset$}
            \STATE Compute $\specvel_{i,t+1}^c$ and $\specpos_{i,t+1}^c$.
	    Examples shown in equations~\eqref{eq:defvcasexn} and~\eqref{eq:defpcasexn}
            \label{alg2:fatitertp1}
            \STATE DISPATCH computation of $\ofunc(\specpos_{i,t+1}^c)$ to a seperate processor.
	    Example shown in equation~\eqref{eq:deffcasexn}.
	    // Note that NONE of the computations, $\specvel_{i,t+1}^c$,
	    $\specpos_{i,t+1}^c$ and $\ofunc(\specpos_{i,t+1}^c)$ depend on $\ofunc(\ppos_{i.t})$.
          \ENDFOR
        \ENDFOR
        \STATE WAIT for all function evaluations dispatched above to complete
        \FOR{$i \in \swarm$}
          \FOR{$c \in \caseset$}
            \STATE Compute $\indic_{i,t+1}^c$
	           Examples shown in equation~\eqref{eq:deficasexn} and~\eqref{eq:defpcasexn}
	  \ENDFOR 
          \STATE Compute $\pvel_{i,t+1}$ as shown in equation~\eqref{eq:vel2update} 
          \STATE Compute $\ppos_{i,t+1}$ as shown in equation~\eqref{eq:pos2update}
          \STATE Compute $\fppos_{i,t+1}$ as shown in equation~\eqref{eq:val2update}
	  \STATE UpdateBests
        \ENDFOR

        \STATE $t=t+1$
        \FOR{$i \in \swarm$}
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE $\nURand_{i,t} \Leftarrow RandomVector()$
          \STATE UpdateBests
          \STATE Update $\pvel_{i,t+1}$ as shown in equation~\eqref{eq:velupdate} 
          \STATE Update $\ppos_{i,t+1}$ as shown in equation~\eqref{eq:posupdate}
        \ENDFOR
    \UNTIL {$t > max \ iteration/2$ or convergence criteria met}
  \end{algorithmic}
\end{algorithm}
